manuscript_figures

### 2024-06-03 scripts/manuscript_figures.Rmd
made figure 3

---
title: "manuscript_figures"
author: "daniellembecker"
date: "2024-05-23"
output: html_document
---

@@

# Load libraries 
```{r}
library(plyr)
library(dplyr)
library(MuMIn)
library(lmerTest)
library(tidyverse)
library(lme4)
library(broom.mixed)
library(lubridate)
library(ggplot2)
library(patchwork)
library(broom)

@@

# Figure 2: Fish biomass (g/m2) on y-axis and Kelp cover (%) on the x-axis. Point scatter plot with a linear regression and 95% CI bands. For all data points, you will have one value for fish biomass and a corresponding value for kelp cover. So it will be by: year x site x transect.

```{r}
# take kelp cover long and just extract data for sugar kelp 2019-2023
kelp.filt <- kelp %>%
        filter(SP_CODE == "SUGK")

@@

# Assuming you want to join by the 'transect' column
merged.dat <- left_join(fish.filt, kelp.filt)

@@

# Make Figure 2
```{r}

# Aggregate the data by year, site, and transect and compute mean values
aggregated_data <- merged.dat %>%
  group_by(YEAR, SITE, TRANSECT) %>%
  summarize(mean_fish_biomass = mean(Fish.Biomass..g.m2., na.rm = TRUE),
            mean_kelp_cover = mean(COVER, na.rm = TRUE)) %>%
  ungroup()

# Fit the linear mixed effects model
lmm_model <- lmer(mean_fish_biomass ~ mean_kelp_cover + SITE + (1 | SITE/YEAR), data = aggregated_data)

@@

# Create scatter plot
figure.2 <- ggplot(aggregated_data, aes(x = mean_kelp_cover, y = mean_fish_biomass)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE, formula = y ~ x, color = "blue") +
annotate("text", x = Inf, y = Inf, label = paste0("Marginal R^2 = ", marginal_r2, "\nConditional R^2 = ", conditional_r2, "\np = ", signif(p_value_kelp_cover, 3)), hjust = 1.4, vjust = 1.1, size = 4, parse = FALSE) + # Adjust vjust
  labs(x = "Kelp cover (%)", y = "Fish biomass (g/m2)",
       title = "Fish Biomass vs Kelp Cover") +
  theme_classic();figure.2

@@

# Figure 3: Make the saame regression figure for fish biomass and kelp biomass (g/m2). To get kelp biomass, you will use the quad data which counts the number of kelp blades, and convert to biomass using the morphometric data. Take stipe length, blade width, blade length and convert to biomass. Then get a mean and SE of kelp frond biomass at that transect and multiply that by the number of fronds counted in quad. We want units to be g/m2 (I think).

```{r}
# read in quad data with filtered for just kelp
quad.filt <- QUAD %>%
        filter(SP_CODE == "SUGK")

@@

#notice some dates and data missing, Austin has to look this over!

#calculate kelp biomass
# To get wet weight in grams (biomass) you use the equation: y=0.00421*x^1.951 where x is kelp blade + stipe length.

# Calculate the kelp blade length
merge.morph.quad$kelp_blade_length <- merge.morph.quad$BLADE_WIDTH_CM * merge.morph.quad$BLADE_LENGTH_CM

# Calculate the sum of blade length and stipe length
merge.morph.quad$sum_length <- merge.morph.quad$kelp_blade_length + merge.morph.quad$STIPE_LENGTH_CM

@@

@@

>>>>

---
title: "manuscript_figures"
author: "daniellembecker"
date: "2024-05-23"
output: html_document
editor_options: 
  chunk_output_type: console
---

@@

# Load libraries 
```{r}
library(plyr)
library(dplyr)
library(car)
library(MuMIn)
library(lmerTest)
library(tidyverse)
library(lme4)
library(broom.mixed)
library(lubridate)
library(ggplot2)
library(patchwork)
library(broom)

@@

# Figure 2: organize data to make fish biomass (g/m2) on y-axis and Kelp cover (%) figure
```{r}
# take kelp cover long and just extract data for sugar kelp 2019-2023
kelp.filt <- kelp %>%
        filter(SP_CODE == "SUGK")

@@

# Assuming you want to join by the 'transect' column
merged.dat.cover <- left_join(fish.filt, kelp.filt)

@@

# Make Figure 2 and run statistical analysis
```{r}
# Aggregate the data by year, site, and transect and compute mean values
aggregated_data_cover <- merged.dat %>%
  group_by(YEAR, SITE, TRANSECT) %>%
  summarize(mean_fish_biomass = mean(Fish.Biomass..g.m2., na.rm = TRUE),
            mean_kelp_cover = mean(COVER, na.rm = TRUE)) %>%
  ungroup()

# Filter dataframe for six top fish biomass above 150
filt_data <- aggregated_data_cover %>%
  filter(mean_fish_biomass > 150)

# Make biomass a percent so * 100
aggregated_data_cover$mean_kelp_cover <- aggregated_data_cover$mean_kelp_cover * 100

# Fit the linear mixed effects model
lmm_model <- lmer(mean_fish_biomass ~ mean_kelp_cover + SITE + (1 | SITE/YEAR), data = aggregated_data_cover)
#check normality
qqPlot(residuals(lmm_model)) #slightly non normal values, log transform to improve

# Fit the linear mixed effects model with log transformation
lmm_model <- lmer(log(mean_fish_biomass) ~ mean_kelp_cover + SITE + (1 | SITE/YEAR), data = aggregated_data_cover)
#check normality
qqPlot(residuals(lmm_model)) #improved to normal

@@

# Create scatter plot
figure.2 <- ggplot(aggregated_data_cover, aes(x = mean_kelp_cover, y = mean_fish_biomass)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE, formula = y ~ x, color = "blue") +
annotate("text", x = Inf, y = Inf, label = paste0("Marginal R^2 = ", marginal_r2, "\nConditional R^2 = ", conditional_r2, "\np = ", signif(p_value_kelp_cover, 3)), hjust = 1.4, vjust = 1.1, size = 4, parse = FALSE) + # Adjust vjust
  labs(x = "Kelp cover (%)", y = "Fish biomass (g/m2)",
       title = "Fish Biomass vs Kelp Cover") +
  theme_classic();figure.2

@@

# Figure 3: organize and calculate kelp biomass data to make fish biomass/kelp biomass (g/m2) linear regression figure
```{r}
# read in quad data with filtered for just kelp
quad.filt <- QUAD %>%
        filter(SP_CODE == "SUGK")

@@

#calculate kelp biomass
# To get wet weight in grams (biomass) you use the equation: y=0.00421*x^1.951 where x is kelp blade + stipe length.
# Calculate the sum of blade length and stipe length
merge.morph.quad$sum_length <- merge.morph.quad$BLADE_LENGTH_CM + merge.morph.quad$STIPE_LENGTH_CM

@@

# Replace NA values with 0
merge.morph.quad$kelp_biomass <- ifelse(is.na(merge.morph.quad$kelp_biomass), 0, merge.morph.quad$kelp_biomass)

# Sum kelp biomass at the transect level
morph.filt <- merge.morph.quad %>%
  group_by(TRANSECT, YEAR, SITE) %>%
  summarize(total_kelp_biomass = sum(kelp_biomass))

# Divide by 6 quadrats for correct g/m2 calculation
morph.filt$average_kelp_biomass <- morph.filt$total_kelp_biomass / 6

@@

# Make figure 3 and run statistical analysis
```{r}
# Aggregate the data by year, site, and transect and compute mean values
merged.dat.morph <- left_join(morph.filt, fish.filt)

# Aggregate the data by year, site, and transect and compute mean values
aggregated.data.morph <- merged.dat.morph %>%
  group_by(YEAR, SITE, TRANSECT) %>%
  summarize(mean_fish_biomass = mean(Fish.Biomass..g.m2., na.rm = TRUE),
            mean_kelp_biomass = mean(average_kelp_biomass, na.rm = TRUE)) %>%
  ungroup()

# Fit the linear mixed effects model
lmm_model <- lmer(mean_fish_biomass ~ mean_kelp_biomass + SITE + (1 | SITE/YEAR), data = aggregated.data.morph)
#check normality
qqPlot(residuals(lmm_model)) #slightly non normal values, log transform to improve

# Fit the linear mixed effects model with log transformation
lmm_model <- lmer(log(mean_fish_biomass) ~ mean_kelp_biomass + SITE + (1 | SITE/YEAR), data = aggregated.data.morph)
#check normality
qqPlot(residuals(lmm_model)) #improved to normal

# Extract p-value for mean_kelp_cover
p_value_kelp_cover <- summary(lmm_model)$coefficients["mean_kelp_biomass", "Pr(>|t|)"]

# Create scatter plot
figure.3 <- ggplot(aggregated.data.morph, aes(x = mean_kelp_biomass, y = mean_fish_biomass)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE, formula = y ~ x, color = "blue") +
annotate("text", x = Inf, y = Inf, label = paste0("Marginal R^2 = ", marginal_r2, "\nConditional R^2 = ", conditional_r2, "\np = ", signif(p_value_kelp_cover, 3)), hjust = 1.4, vjust = 1.1, size = 4, parse = FALSE) + # Adjust vjust
  labs(x = "Kelp Biomass (g/m2)", y = "Fish biomass (g/m2)",
       title = "Fish Biomass vs Kelp Biomass") +
  theme_classic();figure.3

# Save the plot to a file
ggsave("output/figures/figure_3.png", plot = figure.3, width = 8, height = 6)
```
